``Interestingness'' in Mathematical Theory Exploration

### 2019-10-18 PhDThesis/clustering.tex
Fixed todos

\begin{itemize}
\item For simplicity, we use the ``rule of thumb'' given in
  \cite[pp. 365]{mardia1979multivariate} to fix the number of clusters at
  $k = \lceil \sqrt{\frac{n}{2}} \rceil$.
\item Data points will be $d$-dimensional feature vectors, as defined above in
  \S~\ref{sec:featureextraction}.
\item We will use euclidean distance (denoted $e$) as our similarity metric.
\item We will use \emph{k-means} clustering, implemented by Lloyd's algorithm
  \cite{lloyd1982least}.
\end{itemize}

\iffalse
TODO: Alison: It would be good to have justification of the previous
items as well.
\fi

We choose k-means because it is a standard approach \iffalse TODO: Alison: Is
there a reference for this? \fi. There are many other clustering algorithms we
could use, such as \emph{expectation maximisation}, but experiments with
\textsc{ML4PG} have shown little difference in their results; \iffalse TODO:
Alison: Whose experiments? Yours (include something in appendix?) Someone else's
(reference)? MY THOUGHTS: Conversation with Katya; maybe check ML4PG papers, or
omit \fi it seems that \iffalse TODO: Alison: a bit informal and vague \fi the
quality of the features is the bottleneck to learning, so there is no reason to
avoid a fast algorithm like k-means. We use a Haskell implementation provided by
the package \hPackage{k-means}.

Since k-means is iterative, we will use function notation to denote time steps,
so $x(t)$ denotes the value of $x$ at time $t$. \iffalse TODO: Alison: Pick a
tense and stick to it. I'd suggest the present tense throughout \fi We denote
the clusters as $C^1$
to $C^k$. As the name suggests, k-means uses the mean value of each cluster,
which we denote as $\vect{m}^1$ to $\vect{m}^k$. The components of these vectors
are treated as orthogonal, hence the $j$th component of $\vect{m}^i$ is the
average of the $j$th components of all vectors in cluster $C^i$:

@@

Before k-means starts, we must choose \emph{seed} values for
$\vect{m}^i(0)$. Many methods have been proposed for choosing these
values~\cite{arthur2007k}. For simplicity, we will choose values randomly from
our data set $S$; this is known as the Forgy method. \iffalse TODO: Alison: Any
stronger justification? Why are the other methods not appropriate? \fi

@@

Many conservative improvements can be made to the standard k-means algorithm
described above. For example, a more efficient approach like
\emph{yinyang k-means}~\cite{conf/icml/DingZSMM15} could make larger input sizes
more practical to cluster; the \emph{k-means++}
approach~\cite{arthur2007k, bahmani2012scalable} can be used to more carefully
select the ``seed'' values for the first timestep; and the \emph{x-means}
algorithm~\cite{pelleg2000x} is able to estimate how many clusters to use
(instead of relying on our ``rule of thumb''). However, like the choice of
k-means versus other clustering algorithms, the benefits of more, higher-quality
data seem greater than those offered by such algorithmic tweaking. \iffalse
TODO: Also describe better reasons for avoiding these improvements, i.e. they
are algorithmically complicated and require multiple passes over the input data
\fi

>>>>

\begin{itemize}
\item We will use \emph{k-means} clustering~\cite{macqueen1967some}, implemented
  by Lloyd's algorithm~\cite{lloyd1982least}\footnote{We use a Haskell
    implementation provided by the package \hPackage{k-means}.}, due to its
  simplicity, fast convergence~\cite{bottou1994convergence} and use in prior
  work on recurrent clustering~\cite{journals/corr/abs-1212-3618}.
\item The number of clusters $k$ must be chosen (or, if we were to use a
  hierarchical clustering algorithm, a level to ``cut off'' the hierarchy).
  There is no ``correct'' answer, although algorithms like
  \emph{x-means}~\cite{pelleg2000x} can estimate this during clustering. We
  stick to a standard k-means algorithm for the reasons given above, and use
  the common ``rule of thumb''~\cite[pp. 365]{mardia1979multivariate} of setting
  $k = \lceil \sqrt{\frac{n}{2}} \rceil$
\item Data points will be $d$-dimensional feature vectors, as defined above in
  \S~\ref{sec:featureextraction}. This is the standard input format for such
  machine learning tasks.
\item We will use the standard euclidean distance (denoted $e$) as the
  similarity metric between feature vectors. In principle we could construct a
  task-specific distance function to better determine similarity, but that would
  only shift the problem, since it is precisely the task that we are attempting
  to solve through the use of (recurrent) clustering.
\end{itemize}

Since k-means is iterative, we use function notation to denote time steps, so
$x(t)$ denotes the value of $x$ at time $t$. We denote the clusters as $C^1$
to $C^k$. As the name suggests, k-means uses the mean value of each cluster,
which we denote as $\vect{m}^1$ to $\vect{m}^k$. The components of these vectors
are treated as orthogonal, hence the $j$th component of $\vect{m}^i$ is the
average of the $j$th components of all vectors in cluster $C^i$:

@@

Before k-means starts, we must choose \emph{seed} values for
$\vect{m}^i(0)$. Many methods have been proposed for choosing these
values~\cite{arthur2007k}. It is common to choose values randomly from
our data set $S$ (known as the Forgy method), since this prevents having to
synthesise (potentially unrealistic) data.

@@

Many conservative improvements can be made to the standard k-means algorithm
described above, such as the more efficient \emph{yinyang
  k-means}~\cite{conf/icml/DingZSMM15}, or the \emph{k-means++}
approach~\cite{arthur2007k, bahmani2012scalable} to more carefully selecting the
``seed'' values for the first timestep. Adding such complexities does not appear
necessary for the moment, as clustering data sets like our Theory Exploration
Benchmark only takes seconds.
